%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%% copyright.tex LiDIA documentation
%%
%% This file contains LiDIA's copyright
%%
%% Copyright (c) 1995 by the LiDIA Group
%%
%% Authors: Thomas Papanikolaou
%%

\chapter*{Copyright}\addcontentsline{toc}{chapter}{Copyright}
\label{chap:copyright}\markboth{Copyright}{Copyright}

The copyright for \LiDIA is owned by the \LiDIA-Group.  \LiDIA is not in the public domain but
it can be copied and distributed freely for any non-commercial purpose.  Commercial licenses are
available from the \LiDIA-Group.

If you copy \LiDIA for somebody else, you may ask this person for refund of your expenses.  This
should cover cost of media, copying and shipping.  You are not allowed to ask for more than this.
In any case you must give a copy of this copyright notice along with the program.

If you use \LiDIA it is quite a good idea to subscribe to the mailing lists
\url{LiDIA-announce@cdc.informatik.tu-darmstadt.de} to get any \LiDIA-related news, and to
\url{LiDIA@cdc.informatik.tu-darmstadt.de} for discussions on \LiDIA-related topics (see
\url{http://www.cdc.informatik.tu-darmstadt.de/mailman/listinfo/} on how to subscribe).

If you publish a mathematical result that was partly obtained using \LiDIA please cite \LiDIA
(\LiDIA-Group, \LiDIA\ --- \emph{A library for computational number theory}, TU Darmstadt,
2001.), just as you would cite another paper that you used.  Also we would appreciate it if you
could inform us about such a paper (send e-mail to
\url{lidiaadm@cdc.informatik.tu-darmstadt.de}).

You are permitted to modify and redistribute \LiDIA, but you are not allowed to restrict further
redistribution.  That is to say proprietary modifications will not be allowed.  We want all
versions of \LiDIA to remain free.

If you modify any part of \LiDIA and redistribute it, you must supply a README file containing
information about the changes you made.  We do not want to take credit or be blamed for your
modifications.

Of course we are interested in all of your modifications.  In particular we would like to see
bug-fixes, improvements and new functions.  So again we would appreciate it if you informed us
about all modifications you make.

\LiDIA is distributed by us without any warranty, to the extent permitted by applicable state
law.  We distribute \LiDIA \emph{as is} without warranty of any kind, either expressed or
implied, including, but not limited to, the implied warranties of merchantability and fitness
for a particular purpose.

The entire risk as to the quality and performance of the program is with you.  Should \LiDIA
prove defective, you have to take over the cost of all necessary servicing, repair or
correction.

In no case unless required by applicable law will we, and/or any other party who may modify and
redistribute \LiDIA as permitted above, be liable to you for damages, including lost profits,
lost money or other special, incidental or consequential damages arising out of the use or
inability to use \LiDIA.
